Influence bias in cultural evolutionary dynamics

### 2023-11-06 ms/ms.tex
update

@@

We ran simulations of the role-model choice process during a single generation in which every copier evaluates its own optimal success-bias weight, $\alpha^*$, which minimizes the expected squared error between the chosen trait value and the optimal trait value $\hat{A}$,
\begin{equation}\label{eq:optimal_alpha}
\alpha^* = \emph{argmin} \sum_{j=1}^N\frac{\alpha A_j + (1-\alpha) K_j}{\sum_{l=1}^N\alpha A_l + (1-\alpha) K_l} (\hat{A}-A_j)^2 \;,
\end{equation}
where $A_j$ is the trait of role-model $j$ and $K_j$ the number of copiers that already chose role-model~$j$.
Indeed, if the trait value would have been correlated with fitness, the optimal success-bias weight would increase fitness. However in here we don't include natural selection.

@@

@@

>>>>

Similarly to {Henrich and Gil-White} \citep{prestige_evolution}, we believe that the most skilled/knowledgeable models will, on-average, end up with the biggest and most lavish clienteles, so the clientele size provides a convenient and reliable proxy for that person's "worth. Therefore, we predict that skilled individuals have higher status, people preferentially copy them more than others, therefore prestigious individuals are influential, even beyond their domain of expertise.

@@

We ran simulations of the role-model choice process during a single generation in which every copier evaluates its own optimal success-bias weight, $\alpha^*$, which minimizes the expected squared error between the chosen trait value and the optimal trait value $\hat{A}$,
\begin{equation}\label{eq:optimal_alpha}
\alpha^* = \emph{argmin} \sum_{j=1}^N\frac{\alpha A_j + (1-\alpha) K_j}{\sum_{l=1}^N\alpha A_l + (1-\alpha) K_l} (\hat{A}-A_j)^2 \;,
\end{equation}
where $A_j$ is the trait of role-model $j$ and $K_j$ the number of copiers that already chose role-model~$j$.

Simply put, each copier estimates what success-bias weight $\alpha^*$ will result in copying the closet trait to the ideal one $\hat{A}$.
Indeed, if the trait value would have been correlated with fitness, the optimal success-bias weight would increase fitness. However in here we don't include natural selection.

@@

Our decision to research and analyze a multi-bias model is preceded by several others.
{Ze Hong} (2020) \citep{payoff}, for example, present a model that combines both frequency dependent bias (including conformist bias) and payoff bias in a single decision-making calculus. He shows that an integrated learning strategy may be superior to relying on either bias alone. He also show that both are likely to contribute to the spread of variants with high payoff, and the magnitude of conformist bias affects the evolutionary dynamics.

@@

\bibitem{payoff} Hong, Ze. "Combining Conformist and Payoff Bias in Cultural Evolution: An Integrated Model for Human Decision-Making." Human Nature 33.4 (2022): 463-484.
